---
title: "SMDE_assignment03"
author: "Asaf Badouh, Pau Rodriguez"
date: "December 11, 2017"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("rlist")
library(rlist)

sigma = 1.4286
rho = c(0.4, 0.7 , 0.85, 0.925)
E_tau = 77
E_sigma = 15

```
## getMu - explained in the lab assignment paper.
## Allen-Cuneen approximation's formula for G/G/1
## base on miri-32-eng.pdf slide 7
```{r}
getMu <- function(rho, E_tau, sigma){
  m = (E_tau * rho)/(sqrt(exp(sigma*sigma)))
  mu = log(m)
  return(mu)
}

C_s_theta <- function (lamda, mu, rho){
  numerator = (lamda/mu)/(1-rho)
  denominator = 1 + numerator
  return(numerator/denominator)
}

#C(s=1, theta)
allenCuneen <- function (lamda, sigma_tau, mu, sigma_x, rho) {
  C_theta = C_s_theta(lamda, mu, rho)
  #x to the power of y -> x**y
  numerator = lamda**2*sigma_tau**2 + mu**2*sigma_x**2
  numerator = numerator * C_theta
  denominator = 2 * mu * (1-rho)
  
  return(numerator/denominator)
}
```

```{r}
runningQueue <- function(mu, sigma, clients=1000, myseed = -1) {

  
if (myseed > -1){
  set.seed(myseed)
}
service_time = rlnorm(clients,mu,sigma)
inter_arrival = rnorm(clients,77,15) #(t_i) - entrance time instant to the W.S.; it can be obtained through the sequence of interarrival times
L = 0
Lq = 0
W = 0
Wq = 0
LT_i = array(0, clients)
L_i   = array(0, clients)
W_i= array(0, clients)
L_q = array(0, clients)
W_q = array(0, clients)
t_i = array(0, clients)
ts_i = array(0, clients)    #(t_i)^S= arrival time instant to the service system
theta_i = array(0, clients) #Theta_i - exit time instant from W.S. . for client i
for (c in 1:clients){
  #step #1
  if(c == 1)
    ts_i[c] = max(t_i[c])
  else
    ts_i[c] = max(theta_i[c-1], t_i[c])
  #step #2 - done in the pre-processing phase "service_time"
  #step #3
  theta_i[c] = ts_i[c] + service_time[c]
  #step #4
  if(c < clients)
    t_i[c+1] = t_i[c] + inter_arrival[c]
  #step 5 : calculation and printing
  #5a
  L_i[c] = W_i[c] = theta_i[c] - t_i[c]
  L = L + L_i[c]
  #TODO: check what about LT_i[1]
  if (c > 1) 
    LT_i[c] = L/(t_i[c]-t_i[1])
  W = W + W_i[c]
  
  #5b
  L_q[c] = W_q[c] = ts_i[c] - t_i[c]
  Lq = Lq + L_q[c]
  Wq = Wq + W_q[c]
  
}
  
W = W/clients
Wq = Wq/clients
L = L/(t_i[clients] - t_i[1])
Lq = Lq/(t_i[clients] - t_i[1])

 # plot(t_i, LT_i, type="l", xlab="t_i", ylab="LT_i")
  res = list("t_i" = t_i, "LT_i" = LT_i, "W" = W, "Wq"= Wq, "L" = L, "Lq" = Lq)
  res
}

```

```{r}
get_X_mu <- function(rho, X_sigma, Tau_mu, Tau_sigma) {
  
  x_mu = log(rho*Tau_mu/exp((X_sigma**2)/2))
  
  return(x_mu)
  
}

part1_2 <- function(rho, X_sigma, Tau_mu, Tau_sigma){
  
    x_mu = get_X_mu(rho, X_sigma, Tau_mu, Tau_sigma)
    
    C_x = sqrt(exp(X_sigma**2) - 1)
    
    C_tau = Tau_sigma/Tau_mu
    
    Lq_mm1 = (rho**2)/(1 - rho)
    
    Lq_AllenCuneen = Lq_mm1 * (( C_x**2 + C_tau**2 ) /2)
    
    lambda = 1/Tau_mu
    
    Wq_AllenCuneen = Lq_AllenCuneen/lambda
    
    return( c(Lq_AllenCuneen, Wq_AllenCuneen, x_mu))
    
}

```





```{r}
clients = 200000
myseeds = c(7,13,109,211,273,711,777,1001,7001,99)
stat = list()
iterations = 5
for (p in 1:length(rho)){
  mu = getMu(rho[p], E_tau, sigma)
  print(mu)
  for(i in 1:iterations){
    stat = list.append(stat, runningQueue(mu, sigma,clients, myseeds[p %% 10 + 1]))
  }
}

# stat contain all the return values from running Queue where the indexing is:
# stat[1-iterations] results for rho[1] and the corrisponding mu
# stat[iteration+1 - 2*iteration] results for rho[2] and the corrisponding mu
# stat[2*iteration+1 - 3*iteration] results for rho[3] and the corrisponding mu
# stat[3*iteration+1 - 4*iteration] results for rho[4] and the corrisponding mu

for (p in 1:length(rho)){
  for(i in 1:iterations){
    #need to do better plotting, 
    #maybe to plot all the iterations at with the same rho and mu on a same graph.
    plot(stat[i+iterations*(p-1)][[1]]$t_i, stat[i+iterations*(p-1)][[1]]$LT_i, type="l", xlab="t_i", ylab="LT_i")
  }
}


```



## Initial analysis

```{r}

rho07 = read.table('./part1/rho_0_7/queue-sim-arribades.txt')
colnames(rho07)
rho07$V1[3]
# incomplete!
```

## Allen Cuneen's aproximation formula for $W_{q}$ and $L_{q}$

For each loading factor $\rho$, we derive the required $\mu$ value for the Lognormal distribution:
$$
  s = 1 \\
  \lambda = { 1 \over E[\tau] } \\
  \mu = {1 \over E[x]} \\
  E[x] = m \cdot e^{\sigma^{2} \over 2} = e^{\mu + {\sigma^{2} \over 2} } \\
  \rho = { \lambda \over {s\mu}} = { E[x] \over E[\tau] } = e^{\mu} \cdot { e^{\sigma^{2} \over 2} \over E[x] } \Rightarrow \mu = ln \left( { \rho \over e^{ \sigma^{2} \over 2 } } \cdot E[\tau]  \right) \\
$$

We use the Allen Cuneen's aproximation formula for $L_{q}$:
$$
  L_{q} \approx L_{q_{M/M/1}} \cdot \left( C_{\tau}^{2} + C_{x}^{2} \over 2  \right) \\
$$
With:
$$
C_{x} = \sqrt{ \omega - 1} =  \sqrt{ e^{\sigma^{2}} - 1} \\
C_{\tau} = { \sigma_{\tau} \over E[\tau] }
$$
And derive $W_{q}$:
$$
  W_{q} = { L_{q} \over \lambda }
$$

```{r show=FALSE}
# part 1.2 Allen Cuneen aproximation calculation
aproximations = rep(c(0,0), 4)
i = 1
for (r in rho) {
  
  Lq_and_Wq_and_mu = part1_2(r, sigma, E_tau, E_sigma)
  
  aproximations[i] = Lq_and_Wq_and_mu[1]
  aproximations[i+1] = Lq_and_Wq_and_mu[2]
  aproximations[i+2] = Lq_and_Wq_and_mu[3]
  i = i + 3
}
```

Using the Allen Cuneen's aproximation formula, we can compute the $W_{q}$ and $L_{q}$ for each loading factor:

$\rho$ | $\mu$ | $W_{q}$ | $L_{q}$
|-------|---------|---------|--------|
`r rho[1]` | `r aproximations[3]`  | `r aproximations[2]` | `r aproximations[1]` |
`r rho[2]` | `r aproximations[6]`  | `r aproximations[5]` | `r aproximations[4]` |
`r rho[3]` | `r aproximations[9]`  | `r aproximations[8]` | `r aproximations[7]` |
`r rho[4]` |`r aproximations[12]`  | `r aproximations[11]` | `r aproximations[10]` |





## Simulation

First, for each $\rho$, we're going to calculate what is the amount of clients needed to get in the steady state of the waiting system.


```{r  echo=FALSE}
# steady state verification
clients = 100000
stat = list()
iterations = 10
for (p in 1:4){
  mu = getMu(rho[p], E_tau, sigma)
  for(i in 1:1){
    stat = list.append(stat, runningQueue(mu, sigma,clients))
  }
}

for (p in 1:4){
  for(i in 1:1){
    #need to do better plotting, 
    #maybe to plot all the iterations at with the same rho and mu on a same graph.
    plot(stat[[p]]$t_i, stat[[p]]$LT_i, type="l", xlab="t_i", ylab="LT_i")
    title(paste(" rho = ",rho[p]))
  }
}

```

We observe that, apart from the simulation with loading factor 0.4, the other simulations have not attained the steady state.

If we repeat the simulations with a number of clients between 200000 and 500000, the steady state is attained with all loading factors. We have not tested more than 500000 clients.

```{r echo=FALSE}
# steady state verification
clients = 500000
stat = list()
iterations = 10
myseed = c("771","10101", "960", "1078")
for (p in 1:4){
  mu = getMu(rho[p], E_tau, sigma)
  for(i in 1:1){
    stat = list.append(stat, runningQueue(mu, sigma,clients))
  }
}

for (p in 1:4){
  for(i in 1:1){
    #need to do better plotting, 
    #maybe to plot all the iterations at with the same rho and mu on a same graph.
    plot(stat[[p]]$t_i, stat[[p]]$LT_i, type="l", xlab="t_i", ylab="LT_i")
    title(paste(" rho = ",rho[p]))
  }
}

```


### $\rho$ = `r rho[1]`

We generate 10 simulations with 100000 clients, each with a different seed. For each simulation, we check if the steady state is attained without any abrupt increase or decrease in the value of the average occupancy. In case there's any abrupt increase or decrease, we change the seed. If for many seeds, this phenomena is still happening, we increase the number of clients.

```{r echo=FALSE}
simulation <- function(clients, iterations, rho, seeds){
  stat = list()
  mu = getMu(rho, E_tau, sigma)
  for(i in 1:iterations){
    stat = list.append(stat, runningQueue(mu, sigma,clients, seeds[i]))
  }
  
  for(i in 1:iterations){
    #need to do better plotting, 
    #maybe to plot all the iterations at with the same rho and mu on a same graph.
    plot(stat[[i]]$t_i, stat[[i]]$LT_i, type="l", xlab="t_i", ylab="LT_i")
    title(paste(" rho = ",rho," seed = ",seeds[i]))
  }
  
  
  return(stat)
}

clients = 100000
iterations = 10
myseed = c(771,10101, 960, 1078,999,51,89,2001,30719,17)
stat <- simulation(clients,iterations,rho[1], myseeds)

```

We observe that for the seeds  10101, 1078, 960 and 51, there's an abrupt change in the average occupancy. We change those seeds and redo the simulation. 

```{r echo=FALSE}

myseeds = c(771,10102, 963, 1079,999,48,89,2001,30719,17)
stat <- simulation(clients,iterations,rho[1], myseeds)
```


We changed a total of 4 out of 10 seeds. 

We now compute the confidence interval for $L_{q}$ and $W_{q}$. We will use a t-Student distribution with critical value 1 - 0.95 and 9 degrees of freedom to compute a 95% confidence interval.

$$
L_{q_{1}},L_{q_{2}},...,L_{q_{10}} \\
\overline{L}_{q} = {1 \over n}\sum_{i=1}^{n}L_{q_{i}} \\
S_{L_{q}}^{2} = {1 \over n - 1} \sum_{i=1}^{n} ( L_{q_{i}} - \overline{L}_{q} )^{2} \\
C.I.(L_{q}) =   \overline{L}_{q} \pm t_{1 - \alpha, n - 1} \cdot \sqrt{S_{L_{q}}^{2} \over n} \\
$$

$$
W_{q_{1}},W_{q_{2}},...,W_{q_{10}} \\
\overline{W}_{q} = {1 \over n}\sum_{i=1}^{n}W_{q_{i}} \\
S_{W_{q}}^{2} = {1 \over n - 1} \sum_{i=1}^{n} ( W_{q_{i}} - \overline{W}_{q} )^{2} \\
C.I.(W_{q}) =   \overline{W}_{q} \pm t_{1 - \alpha, n - 1} \cdot \sqrt{S_{W_{q}}^{2} \over n} \\
$$


The computations produce the followings confidence intervals for the average queue length and waiting time:

```{r echo=FALSE}

confidence_interval <- function(x,r) {

  n = length(x)
  # gather all Lq's and Wq's
  vLq <- c()
  vWq <- c()
  for (i in 1:n){
    vLq[i] <- x[[i]]$Lq
    vWq[i] <- x[[i]]$Wq
  }
  
  # compute the mean
  Lqavg = mean(vLq)
  Wqavg = mean(vWq)
  
  # compute the sample variance
  
  Lqs2 = sum((vLq - Lqavg)**2)/(n - 1)
  Wqs2 = sum((vWq - Wqavg)**2)/(n - 1)
  
  # t-student
  t = qt(0.95,n - 1)
  
  # confidence interval
  CI_Lq = c(Lqavg - t*sqrt(Lqs2/n), Lqavg + t*sqrt(Lqs2/n) )
  CI_Wq = c(Wqavg - t*sqrt(Wqs2/n), Wqavg + t*sqrt(Wqs2/n) )

  return(data.frame(rho=r,Lq_m=CI_Lq[1], Lq_p=CI_Lq[2], Wq_m=CI_Wq[1], Wq_p=CI_Wq[2] ))  
}

ci <- data.frame(rho=c(),lq_m=c(),lq_p=c(),wq_m=c(),wq_p=c())
ci_current <- confidence_interval(stat,rho[1])
ci <- rbind(ci, setNames(ci_current, names(ci)))
ci[1,]
```


### $\rho$ = `r rho[2]`

We generate 10 simulations with 100000 clients, each with a different seed. We check if the steady state is attained. We also check for irregularities in the simulation results, in which case we change the random number generator seed and repeat the simulation.

```{r echo=FALSE}

clients = 100000
iterations = 10
myseeds = c(771,10101, 960, 1078,999,51,89,2001,30719,17)
myseeds = c(772,10102, 963, 1078,999,48,89,2001,30718,17)
stat <- simulation(clients,iterations,rho[2], myseeds)

```

We observe that 5 out of 10 simulations are not in a steady state at the end of the simulation. We increase the number of clients to 500000 and repeat the simulations.

```{r echo=FALSE}

clients = 500000
iterations = 10
myseeds = c(772,10102, 963, 1078,999,48,89,2001,30718,17)
stat <- simulation(clients,iterations,rho[2], myseeds)

```

We now compute the confidence interval for $L_{q}$ and $W_{q}$ using the same procedure we detailed for $\rho =$ `r rho[1]`. We will use a t-Student distribution with critical value 1 - 0.95 and 9 degrees of freedom to compute a 95% confidence interval. The computations produce the followings confidence intervals for the average queue length and waiting time:

```{r echo=FALSE}
ci_current <- confidence_interval(stat,rho[2])
ci <- rbind(ci, setNames(ci_current, names(ci)))
ci[2,]
```

### $\rho$ = `r rho[3]`

We generate 10 simulations with 100000 clients, each with a different seed. We check if the steady state is attained. We also check for irregularities in the simulation results, in which case we change the random number generator seed and repeat the simulation.

```{r echo=FALSE}

clients = 100000
iterations = 10
myseeds = c(771,10101, 960, 1078,999,51,89,2001,30719,17)
myseeds = c(772,10102, 963, 1078,999,48,89,2001,30718,17)
stat <- simulation(clients,iterations,rho[3], myseeds)

```

We observe that ... PENDING 

We now compute the confidence interval for $L_{q}$ and $W_{q}$ using the same procedure we detailed for $\rho =$ `r rho[1]`. We will use a t-Student distribution with critical value 1 - 0.95 and 9 degrees of freedom to compute a 95% confidence interval. The computations produce the followings confidence intervals for the average queue length and waiting time:

```{r echo=FALSE}
ci_current <- confidence_interval(stat,rho[3])
ci <- rbind(ci, setNames(ci_current, names(ci)))
ci[3,]
```

### $\rho$ = `r rho[4]`

We generate 10 simulations with 100000 clients, each with a different seed. We check if the steady state is attained. We also check for irregularities in the simulation results, in which case we change the random number generator seed and repeat the simulation.

```{r echo=FALSE}

clients = 100000
iterations = 10
myseeds = c(771,10101, 960, 1078,999,51,89,2001,30719,17)
myseeds = c(772,10102, 963, 1078,999,48,89,2001,30718,17)
stat <- simulation(clients,iterations,rho[4], myseeds)

```

We observe that ... PENDING 

We now compute the confidence interval for $L_{q}$ and $W_{q}$ using the same procedure we detailed for $\rho =$ `r rho[1]`. We will use a t-Student distribution with critical value 1 - 0.95 and 9 degrees of freedom to compute a 95% confidence interval. The computations produce the followings confidence intervals for the average queue length and waiting time:

```{r echo=FALSE}
ci_current <- confidence_interval(stat,rho[4])
ci <- rbind(ci, setNames(ci_current, names(ci)))
ci[4,]
```

## Comparison of Allen Cuneen's approximation and the simulation

$\rho$ | $W_{q}$ | -C.I $W_{q}$ | +C.I $W_{q}$ | $L_{q}$ | -C.I $L_{q}$ | +C.I $L_{q}$ |
|-------|---------|---------|--------|---|---|---|---|
`r rho[1]` |  `r aproximations[2]` | `r ci[1,4] ` | `r ci[1,5] ` | `r aproximations[1]` | `r ci[1,2] ` | `r ci[1,3] `
`r rho[2]` |  `r aproximations[5]` | `r ci[2,4] ` | `r ci[2,5] ` | `r aproximations[4]` | `r ci[2,2] ` | `r ci[2,3] `
`r rho[3]` | `r aproximations[8]` | `r ci[3,4] ` | `r ci[3,5] ` | `r aproximations[7]` | `r ci[3,2] ` | `r ci[3,3] `
`r rho[4]` | `r aproximations[11]` | `r ci[4,4] ` | `r ci[4,5] ` | `r aproximations[10]` | `r ci[4,2] ` | `r ci[4,3] `


