---
title: "SMDE - Assignment 03"
author: "Asaf Badouh, Pau Rodriguez"
date: "January 12, 2018"
output:
  pdf_document:
    keep_tex: true
    fig_caption: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("rlist")
library(rlist)
knitr::opts_chunk$set(cache=TRUE)



```


```{r echo=FALSE}
# Distribution attributes
sigma = 1.4286
rho = c(0.4, 0.7 , 0.85, 0.925)
E_tau = 77
E_sigma = 15


# getMu - explained in the lab assignment paper.
# Allen-Cuneen approximation's formula for G/G/1
# base on miri-32-eng.pdf slide 7

getMu <- function(rho, E_tau, sigma){
  m = (E_tau * rho)/(sqrt(exp(sigma*sigma)))
  mu = log(m)
  return(mu)
}

C_s_theta <- function (lamda, mu, rho){
  numerator = (lamda/mu)/(1-rho)
  denominator = 1 + numerator
  return(numerator/denominator)
}

#C(s=1, theta)
allenCuneen <- function (lamda, sigma_tau, mu, sigma_x, rho) {
  C_theta = C_s_theta(lamda, mu, rho)
  #x to the power of y -> x**y
  numerator = lamda**2*sigma_tau**2 + mu**2*sigma_x**2
  numerator = numerator * C_theta
  denominator = 2 * mu * (1-rho)
  
  return(numerator/denominator)
}
```

```{r echo=FALSE }
runningQueue <- function(mu, sigma, clients=1000, myseed = -1) {

  
if (myseed > -1){
  set.seed(myseed)
}
service_time = rlnorm(clients,mu,sigma)
inter_arrival = rnorm(clients,77,15) #(t_i) - entrance time instant to the W.S.; it can be obtained through the sequence of interarrival times
L = 0
Lq = 0
W = 0
Wq = 0
LT_i = array(0, clients)
L_i   = array(0, clients)
W_i= array(0, clients)
L_q = array(0, clients)
W_q = array(0, clients)
t_i = array(0, clients)
ts_i = array(0, clients)    #(t_i)^S= arrival time instant to the service system
theta_i = array(0, clients) #Theta_i - exit time instant from W.S. . for client i
for (c in 1:clients){
  #step #1
  if(c == 1)
    ts_i[c] = max(t_i[c])
  else
    ts_i[c] = max(theta_i[c-1], t_i[c])
  #step #2 - done in the pre-processing phase "service_time"
  #step #3
  theta_i[c] = ts_i[c] + service_time[c]
  #step #4
  if(c < clients)
    t_i[c+1] = t_i[c] + inter_arrival[c]
  #step 5 : calculation and printing
  #5a
  L_i[c] = W_i[c] = theta_i[c] - t_i[c]
  L = L + L_i[c]
  #TODO: check what about LT_i[1]
  if (c > 1) 
    LT_i[c] = L/(t_i[c]-t_i[1])
  W = W + W_i[c]
  
  #5b
  L_q[c] = W_q[c] = ts_i[c] - t_i[c]
  Lq = Lq + L_q[c]
  Wq = Wq + W_q[c]
  
}
  
W = W/clients
Wq = Wq/clients
L = L/(t_i[clients] - t_i[1])
Lq = Lq/(t_i[clients] - t_i[1])

 # plot(t_i, LT_i, type="l", xlab="t_i", ylab="LT_i")
  res = list("t_i" = t_i, "LT_i" = LT_i, "W" = W, "Wq"= Wq, "L" = L, "Lq" = Lq, "W_i"=W_i, "W_q"=W_q)
  res
}

```

```{r echo=FALSE}
get_X_mu <- function(rho, X_sigma, Tau_mu, Tau_sigma) {
  
  x_mu = log(rho*Tau_mu/exp((X_sigma**2)/2))
  
  return(x_mu)
  
}

part1_2 <- function(rho, X_sigma, Tau_mu, Tau_sigma){
  
    x_mu = get_X_mu(rho, X_sigma, Tau_mu, Tau_sigma)
    
    C_x = sqrt(exp(X_sigma**2) - 1)
    
    C_tau = Tau_sigma/Tau_mu
    
    Lq_mm1 = (rho**2)/(1 - rho)
    
    Lq_AllenCuneen = Lq_mm1 * (( C_x**2 + C_tau**2 ) /2)
    
    lambda = 1/Tau_mu
    
    Wq_AllenCuneen = Lq_AllenCuneen/lambda
    
    return( c(Lq_AllenCuneen, Wq_AllenCuneen, x_mu))
    
}

```





```{r echo=FALSE }

clients = 200000
myseeds = c(7,13,109,211,273,711,777,1001,7001,99)
stat = list()
iterations = 5

simulation.test <- function (){  
    for (p in 1:length(rho)){
    mu = getMu(rho[p], E_tau, sigma)
    print(mu)
    for(i in 1:iterations){
      stat = list.append(stat, runningQueue(mu, sigma,clients, myseeds[p %% 10 + 1]))
    }
  }
  
  # stat contain all the return values from running Queue where the indexing is:
  # stat[1-iterations] results for rho[1] and the corrisponding mu
  # stat[iteration+1 - 2*iteration] results for rho[2] and the corrisponding mu
  # stat[2*iteration+1 - 3*iteration] results for rho[3] and the corrisponding mu
  # stat[3*iteration+1 - 4*iteration] results for rho[4] and the corrisponding mu
  
  for (p in 1:length(rho)){
    for(i in 1:iterations){
      #need to do better plotting, 
      #maybe to plot all the iterations at with the same rho and mu on a same graph.
      plot(stat[i+iterations*(p-1)][[1]]$t_i, stat[i+iterations*(p-1)][[1]]$LT_i, type="l", xlab="t_i", ylab="LT_i")
    }
  }
  

    
}

```




## Initial analysis

|       |  Interarrival Times $\tau$      | Service times $X$ |
|-----------------------|:---------------:|:---------------:|
| Assigned Distributions | Normal         | Lognormal  | 
|  Parameters | ($\mu=77$, $\sigma=15$)   |  ($\mu$, $\sigma=1.4286$) | 

We execute a small simulation with 10.000 clients to verify that the simulation is running correctly. 
Our probability distribution for modelling the services times is a Lognormal with $\sigma=1.4286$. To obtain the other parameter of the distribution, the $\mu$, we choose the first traffic factor, $\rho=0.4$, which allows us to compute the $\mu$.  
$$  \rho = { \lambda \over {s\mu}} = { E[x] \over E[\tau] } = e^{\mu} \cdot { e^{\sigma^{2} \over 2} \over E[x] } \Rightarrow \mu = ln \left( { \rho \over e^{ \sigma^{2} \over 2 } } \cdot E[\tau]  \right) $$
The resulting $\mu$ is 2.4070657.

In our simulation, the arrival times are defined by a Normal distributioni with parameters $\mu=77, \sigma=15$. We generate 10.000 clients int the simulation, and analyse the service times.
```{r echo=FALSE}
clients = 10000
stat = list()
mu = getMu(rho[1], E_tau, sigma)
stat = list.append(stat, runningQueue(mu, sigma, clients))
Wi = stat[[1]]$"W_i"
Wqi = stat[[1]]$"W_q"
```


The histogram of the service times is shown below:
```{r echo=FALSE,out.width = "50%", fig.align='center'}
hist(Wi - Wqi, breaks=700, probability=TRUE) 
```

```{r echo=FALSE}
# mean
ws_mean = mean(Wi - Wqi)

# std
ws_std = sqrt(sum((ws_mean - (Wi - Wqi))^2)/(length(Wi) - 1) )

# Coefficient of variation
cx = ws_std/ws_mean

# theoretical values
m = exp(mu)
omega = exp(sigma^2)
Ex = m*sqrt(omega)
Varx = m*m*omega*(omega - 1)
stdx = sqrt(Varx)
ctx = stdx/Ex
```

The mean and sample variance of the services times are the following:

1. Mean of the service times: $W_{s_{avg}}$=`r ws_mean `
2. Standard deviation of the service times: $W_{s_{std}}$=`r ws_std`
3. Coeficient of variation: $C_{x}={ \sigma_{x} \over E[x] }$=`r cx`

The Theoretical values for a Lognormal with $\mu$=`r mu ` and $\sigma$=`r sigma` are the follwing:

1. Mean of the service times: $E[x]$=`r Ex`
2. Standard deviation of the service times: $sqrt{Var[x]}$=`r stdx`
3. Coeficient of variation: $C_{x}={ \sigma_{x} \over E[x] }$=`r ctx`

We observe that the sample statistics are close to the theoretical values. Our simulation can be considered to be correct.



```{r echo=FALSE}

#rho07 = read.table('./part1/rho_0_7/queue-sim-arribades.txt')
#colnames(rho07)
#rho07$V1[3]

# incomplete!
```

## Allen Cuneen's aproximation formula for $W_{q}$ and $L_{q}$

For each loading factor $\rho$, we derive the required $\mu$ value for the Lognormal distribution:

- $s = 1$,  $\lambda = { 1 \over E[\tau] }$ , $\mu = {1 \over E[x]}$
- $E[x] = m \cdot e^{\sigma^{2} \over 2} = e^{\mu + {\sigma^{2} \over 2} }$
- $\rho = { \lambda \over {s\mu}} = { E[x] \over E[\tau] } = e^{\mu} \cdot { e^{\sigma^{2} \over 2} \over E[x] } \Rightarrow \mu = ln \left( { \rho \over e^{ \sigma^{2} \over 2 } } \cdot E[\tau]  \right)$

We use the Allen Cuneen's approximation formula for $L_{q}$:

- $L_{q} \approx L_{q_{M/M/1}} \cdot \left( C_{\tau}^{2} + C_{x}^{2} \over 2  \right)$ 
- with: $C_{x} = \sqrt{ \omega - 1} =  \sqrt{ e^{\sigma^{2}} - 1}$  and $C_{\tau} = { \sigma_{\tau} \over E[\tau] }$
- and derive $W_{q} = { L_{q} \over \lambda }$

```{r echo=FALSE}
# part 1.2 Allen Cuneen aproximation calculation
aproximations = rep(c(0,0), 4)
i = 1
for (r in rho) {
  
  Lq_and_Wq_and_mu = part1_2(r, sigma, E_tau, E_sigma)
  
  aproximations[i] = Lq_and_Wq_and_mu[1]
  aproximations[i+1] = Lq_and_Wq_and_mu[2]
  aproximations[i+2] = Lq_and_Wq_and_mu[3]
  i = i + 3
}
```

Using the Allen Cuneen's approximation formula, we can compute the $W_{q}$ and $L_{q}$ for each loading factor:

$\rho$ | $\mu$ | $W_{q}$ | $L_{q}$
|-------|---------|---------|--------|
`r rho[1]` | `r aproximations[3]`  | `r aproximations[2]` | `r aproximations[1]` |
`r rho[2]` | `r aproximations[6]`  | `r aproximations[5]` | `r aproximations[4]` |
`r rho[3]` | `r aproximations[9]`  | `r aproximations[8]` | `r aproximations[7]` |
`r rho[4]` |`r aproximations[12]`  | `r aproximations[11]` | `r aproximations[10]` |





## Simulation

First, for each $\rho$, we're going to calculate what is the amount of clients needed to get in the steady state of the waiting system.


```{r  echo=FALSE}
# steady state verification
clients = 100000
stat = list()
iterations = 10
for (p in 1:4){
  mu = getMu(rho[p], E_tau, sigma)
  for(i in 1:1){
    stat = list.append(stat, runningQueue(mu, sigma,clients))
  }
}

for (p in 1:4){
  for(i in 1:1){
    #need to do better plotting, 
    #maybe to plot all the iterations at with the same rho and mu on a same graph.
    plot(stat[[p]]$t_i, stat[[p]]$LT_i, type="l", xlab="t_i", ylab="LT_i")
    title(paste(" rho = ",rho[p]))
  }
}

```

We observe that, apart from the simulation with loading factor 0.4, the other simulations have not attained the steady state.

If we repeat the simulations with a number of clients between 200000 and 500000, the steady state is attained with all loading factors. We have not tested more than 500000 clients.

```{r echo=FALSE}
# steady state verification
clients = 500000
stat = list()
iterations = 10
myseed = c("771","10101", "960", "1078")
for (p in 1:4){
  mu = getMu(rho[p], E_tau, sigma)
  for(i in 1:1){
    stat = list.append(stat, runningQueue(mu, sigma,clients))
  }
}

for (p in 1:4){
  for(i in 1:1){
    #need to do better plotting, 
    #maybe to plot all the iterations at with the same rho and mu on a same graph.
    plot(stat[[p]]$t_i, stat[[p]]$LT_i, type="l", xlab="t_i", ylab="LT_i")
    title(paste(" rho = ",rho[p]))
  }
}

```


###  Loading factor $\rho$ = `r rho[1]`

We generate 10 simulations with 100000 clients, each with a different seed. For each simulation, we check if the steady state is attained without any abrupt increase or decrease in the value of the average occupancy. In case there's any abrupt increase or decrease, we change the seed. If for many seeds, this phenomena is still happening, we increase the number of clients.

```{r echo=FALSE}
simulation <- function(clients, iterations, rho, seeds){
  stat = list()
  mu = getMu(rho, E_tau, sigma)
  for(i in 1:iterations){
    stat = list.append(stat, runningQueue(mu, sigma,clients, seeds[i]))
  }
  
  for(i in 1:iterations){
    #need to do better plotting, 
    #maybe to plot all the iterations at with the same rho and mu on a same graph.
    plot(stat[[i]]$t_i, stat[[i]]$LT_i, type="l", xlab="t_i", ylab="LT_i")
    title(paste(" rho = ",rho," seed = ",seeds[i]))
  }
  
  
  return(stat)
}
```

```{r echo=FALSE}
clients = 100000
iterations = 10
myseed = c(771,10101, 960, 1078,999,51,89,2001,30719,17)
stat <- simulation(clients,iterations,rho[1], myseeds)

```

We observe that for the seeds  10101, 1078, 960 and 51, there's an abrupt change in the average occupancy. We change those seeds and redo the simulation. 

```{r echo=FALSE}

myseeds = c(771,10102, 963, 1079,999,48,89,2001,30719,17)
stat <- simulation(clients,iterations,rho[1], myseeds)
```


We changed a total of 4 out of 10 seeds. 

We now compute the confidence interval for $L_{q}$ and $W_{q}$. We will use a t-Student distribution with critical value 1 - 0.95 and 9 degrees of freedom to compute a 95% confidence interval.

- $L_{q_{1}},L_{q_{2}},...,L_{q_{10}}$, and $\overline{L}_{q} = {1 \over n}\sum_{i=1}^{n}L_{q_{i}}$
- $S_{L_{q}}^{2} = {1 \over n - 1} \sum_{i=1}^{n} ( L_{q_{i}} - \overline{L}_{q} )^{2}$

To get:
$$C.I.(L_{q}) =   \overline{L}_{q} \pm t_{1 - \alpha, n - 1} \cdot \sqrt{S_{L_{q}}^{2} \over n}$$

And then:

- $W_{q_{1}},W_{q_{2}},...,W_{q_{10}}$,  and $\overline{W}_{q} = {1 \over n}\sum_{i=1}^{n}W_{q_{i}}$
- $S_{W_{q}}^{2} = {1 \over n - 1} \sum_{i=1}^{n} ( W_{q_{i}} - \overline{W}_{q} )^{2}$

to get:
$$ C.I.(W_{q}) =   \overline{W}_{q} \pm t_{1 - \alpha, n - 1} \cdot \sqrt{S_{W_{q}}^{2} \over n}$$

The computations produce the followings confidence intervals for the average queue length and waiting time:

```{r echo=FALSE}

confidence_interval <- function(x,r) {

  n = length(x)
  # gather all Lq's and Wq's
  vLq <- c()
  vWq <- c()
  for (i in 1:n){
    vLq[i] <- x[[i]]$Lq
    vWq[i] <- x[[i]]$Wq
  }
  
  # compute the mean
  Lqavg = mean(vLq)
  Wqavg = mean(vWq)
  
  # compute the sample variance
  
  Lqs2 = sum((vLq - Lqavg)**2)/(n - 1)
  Wqs2 = sum((vWq - Wqavg)**2)/(n - 1)
  
  # t-student
  t = qt(0.95,n - 1)
  
  # confidence interval
  CI_Lq = c(Lqavg - t*sqrt(Lqs2/n), Lqavg + t*sqrt(Lqs2/n) )
  CI_Wq = c(Wqavg - t*sqrt(Wqs2/n), Wqavg + t*sqrt(Wqs2/n) )

  return(data.frame(rho=r,Lq_m=CI_Lq[1], Lq_p=CI_Lq[2], Wq_m=CI_Wq[1], Wq_p=CI_Wq[2] ))  
}

ci <- data.frame(rho=c(),lq_m=c(),lq_p=c(),wq_m=c(),wq_p=c())
ci_current <- confidence_interval(stat,rho[1])
ci <- rbind(ci, setNames(ci_current, names(ci)))
#ci[1,]
```

$\rho$  | -C.I $W_{q}$ | +C.I $W_{q}$ | -C.I $L_{q}$ | +C.I $L_{q}$ |
|-------|---------|---------|--------|---|---|
`r ci[1,1]` | `r ci[1,2]` | `r ci[1,3]` | `r ci[1,4]` | `r ci[1,5]`



### Loading factor $\rho$ = `r rho[2]`

We follow the same procedure to generate 10 simulations with 100000 clients, as we did for the previous $\rho=$, correcting any invalid seeds for the ramdom number generator. The result is that we need to increase the number of clients to 500000 to be sure that the system has arribed to the steady state.

```{r echo=FALSE}

clients = 100000
iterations = 10
myseeds = c(771,10101, 960, 1078,999,51,89,2001,30719,17)
myseeds = c(772,10102, 963, 1078,999,48,89,2001,30718,17)
#stat <- simulation(clients,iterations,rho[2], myseeds)

```


```{r echo=FALSE}
#We observe that 5 out of 10 simulations are not in a steady state at the end of the simulation. We increase the number of clients to 500000 and repeat the simulations.


clients = 500000
iterations = 10
myseeds = c(772,10102, 963, 1078,999,48,89,2001,30718,17)
stat <- simulation(clients,iterations,rho[2], myseeds)

```

We now compute the confidence interval for $L_{q}$ and $W_{q}$ using the same procedure we detailed for $\rho =$ `r rho[2]`. The computations produce the followings confidence intervals for the average queue length and waiting time:

```{r echo=FALSE}
ci_current <- confidence_interval(stat,rho[2])
ci <- rbind(ci, setNames(ci_current, names(ci)))
```

$\rho$  | -C.I $W_{q}$ | +C.I $W_{q}$ | -C.I $L_{q}$ | +C.I $L_{q}$ |
|-------|---------|---------|--------|---|---|
`r ci[2,1]` | `r ci[2,2]` | `r ci[2,3]` | `r ci[2,4]` | `r ci[2,5]`

### Loading factor $\rho$ = `r rho[3]`

We generate 10 simulations with 100000 clients, each with a different seed. We check if the steady state is attained. We also check for irregularities in the simulation results, in which case we change the random number generator seed and repeat the simulation.

```{r echo=FALSE}

clients = 100000
iterations = 10
myseeds = c(771,10101, 960, 1078,999,51,89,2001,30719,17)
myseeds = c(772,10102, 963, 1078,999,48,89,2001,30718,17)
#stat <- simulation(clients,iterations,rho[3], myseeds)
#We observe that seeds 963, 1078, 48 and 2001 produce irregularities in the simulation, therefore we have to change them and repeat the simulation.
```



```{r echo=FALSE}

clients = 100000
iterations = 10
myseeds = c(771,10101, 960, 1078,999,51,89,2001,30719,17)
myseeds = c(772,10102, 963, 1078,999,48,89,2001,30718,17)
myseeds = c(772,10102, 970, 1151,999,55,89,3001,30718,17)
stat <- simulation(clients,iterations,rho[3], myseeds)

```

We now compute the confidence interval for $L_{q}$ and $W_{q}$ using the same procedure we detailed for $\rho =$ `r rho[3]`. The computations produce the followings confidence intervals for the average queue length and waiting time:

```{r echo=FALSE}
ci_current <- confidence_interval(stat,rho[3])
ci <- rbind(ci, setNames(ci_current, names(ci)))

```

$\rho$  | -C.I $W_{q}$ | +C.I $W_{q}$ | -C.I $L_{q}$ | +C.I $L_{q}$ |
|-------|---------|---------|--------|---|---|
`r ci[3,1]` | `r ci[3,2]` | `r ci[3,3]` | `r ci[3,4]` | `r ci[3,5]`

### Loading factor $\rho$ = `r rho[4]`

We generate 10 simulations with 100000 clients, each with a different seed. We check if the steady state is attained. We also check for irregularities in the simulation results, in which case we change the random number generator seed and repeat the simulation.

```{r echo=FALSE}

clients = 100000
iterations = 10
myseeds = c(771,10101, 960, 1078,999,51,89,2001,30719,17)
myseeds = c(772,10102, 963, 1078,999,48,89,2001,30718,17)
myseeds = c(772,10102, 970, 1151,999,55,89,3001,30718,17)
stat <- simulation(clients,iterations,rho[4], myseeds)

```

We now compute the confidence interval for $L_{q}$ and $W_{q}$ using the same procedure we detailed for $\rho =$ `r rho[4]`. The computations produce the followings confidence intervals for the average queue length and waiting time:

```{r echo=FALSE}
ci_current <- confidence_interval(stat,rho[4])
ci <- rbind(ci, setNames(ci_current, names(ci)))

```
$\rho$  | -C.I $W_{q}$ | +C.I $W_{q}$ | -C.I $L_{q}$ | +C.I $L_{q}$ |
|-------|---------|---------|--------|---|---|
`r ci[4,1]` | `r ci[4,2]` | `r ci[4,3]` | `r ci[4,4]` | `r ci[4,5]`



## Comparison of Allen Cuneen's approximation and the simulation

$\rho$ | $W_{q}$ | -C.I $W_{q}$ | +C.I $W_{q}$ | $L_{q}$ | -C.I $L_{q}$ | +C.I $L_{q}$ |
|-------|---------|---------|--------|---|---|---|---|
`r rho[1]` |  `r aproximations[2]` | `r ci[1,4] ` | `r ci[1,5] ` | `r aproximations[1]` | `r ci[1,2] ` | `r ci[1,3] `
`r rho[2]` |  `r aproximations[5]` | `r ci[2,4] ` | `r ci[2,5] ` | `r aproximations[4]` | `r ci[2,2] ` | `r ci[2,3] `
`r rho[3]` | `r aproximations[8]` | `r ci[3,4] ` | `r ci[3,5] ` | `r aproximations[7]` | `r ci[3,2] ` | `r ci[3,3] `
`r rho[4]` | `r aproximations[11]` | `r ci[4,4] ` | `r ci[4,5] ` | `r aproximations[10]` | `r ci[4,2] ` | `r ci[4,3] `

As we can see in the table above Allen-Cuneen's approximation values for all loading factors are outside of the confidence interval of our simulations, always above. Overall, the approximations of both, $W_q$ and $L_{q}$, are not very far from the simulation's confidence interval.

So the Allen-Cuneen's aproximation models a system where the occupancy is higher and the waiting times in the queue are also greater than what our simulation produces. How can we explain that?

First of all, our random number generator has not been tested. As discussed  in class, it is impossible to generate real randomness, so every rng must be tested for multiple desirable properties. In our particular case, even though we used the rng that ships with R, which is considered to ship with well tested generated random number streams, we noticed that while repeatedly running the simulation we needed to change the seed from time to time in order to avoid irregularities or abrupt changes. We can conclude that we need to test the rng for appropriate period, given that we need around 200.000 samples in each simulation.

Secondly, the Allen Cuneen's approximation uses the exact values for exponential models and a correction factor for our model.
Causes to the difference between aproximation and simulation: 

  1. Our model is not so close to a heavy tail distribution, so the correction factor is not accurate in the Allen Cuneen's formula.
  2. We cannot use the Allen Cuneen's approximation formula with our model. We need to use the Kôllerstrôm formula, as our system is in "heavy traffic" conditions (it is a heavy tail distribution and that means heavy traffic?)
  3. Something related to Ctau and Cx given that arrivals is a Normal distr. and service times is lognormal?




```{r echo=FALSE}

#kollerstrom teorem

kWq1 = (1/77)^2 * (77^2  + (1/1) * sigma^2)/((2 * (1 - rho[1])))
kWq2 = (1/77)^2 * (77^2  + (1/1) * sigma^2)/((2 * (1 - rho[2])))
kWq3 = (1/77)^2 * (77^2  + (1/1) * sigma^2)/((2 * (1 - rho[3])))
kWq4 = (1/77)^2 * (77^2  + (1/1) * sigma^2)/((2 * (1 - rho[4])))


kLq1 = ((1/77)^2 * 15^2 + rho[1]^2 * mu^2 * sigma^2) / (2 * (1 - rho[1]))
kLq2 = ((1/77)^2 * 15^2 + rho[2]^2 * mu^2 * sigma^2) / (2 * (1 - rho[2]))
kLq3 = ((1/77)^2 * 15^2 + rho[3]^2 * mu^2 * sigma^2) / (2 * (1 - rho[3]))
kLq4 = ((1/77)^2 * 15^2 + rho[4]^2 * mu^2 * sigma^2) / (2 * (1 - rho[4]))

```


$\rho$ | $W_{q}$ | -C.I $W_{q}$ | +C.I $W_{q}$ | $L_{q}$ | -C.I $L_{q}$ | +C.I $L_{q}$ |
|-------|---------|---------|--------|---|---|---|---|
`r rho[1]` |  `r kWq1` | `r ci[1,4] ` | `r ci[1,5] ` | `r kLq1` | `r ci[1,2] ` | `r ci[1,3] `
`r rho[2]` |  `r kWq2` | `r ci[2,4] ` | `r ci[2,5] ` | `r kLq2` | `r ci[2,2] ` | `r ci[2,3] `
`r rho[3]` | `r kWq3` | `r ci[3,4] ` | `r ci[3,5] ` | `r kLq3` | `r ci[3,2] ` | `r ci[3,3] `
`r rho[4]` | `r kWq4` | `r ci[4,4] ` | `r ci[4,5] ` | `r kLq4` | `r ci[4,2] ` | `r ci[4,3] `

We can observe that Köllerström formula does not apply to this situation. We conclude that we are not in heavy traffic conditions.
